\chapter{Lab Maintenance} \label{ap:lab_maintenance}

\begin{verbatim}

#############################
#!/bin/bash
# resetX

# get the current time (use in naming backup files)
TIME=`date +%y%m%d%H%M` # echo $TIME
#

# save Old Working Directory
OWD="/net/d2/u5/f/conf"

# make sure we are in HOME directory
cd ~

# make a directory to save old conf
SAVE=.oldXconfig.$TIME #echo $SAVE
mkdir -p $SAVE

for FILE in `cat $OWD/Xrc.confiles`  ; do
  if [ -e $FILE ] ; then
     #echo $FILE
     mv -f $FILE $SAVE/
  fi
done
#ls -rlast $SAVE
tar -xf ~conf/data/skel.tar

#/bin/rm -rf /tmp/*$USER* >& /dev/null
#/bin/rm -rf /tmp/.*$USER* >& /dev/null 
/bin/rm -rf /tmp/.* >& /dev/null 
/bin/rm -rf /tmp/* >& /dev/null 

if [ -f ~/Desktop/xterm.desktop ] ; then
	/bin/rm -f ~/Desktop/xterm.desktop
fi
if [ -f ~/Desktop/TerminalServer.desktop ] ; then
	/bin/rm -f ~/Desktop/Terminal\ Server.desktop
fi

/bin/cp -fv ~conf/Desktop/* ~/Desktop/



#############################
#!/bin/sh
# This script takes in a file of student homes, then cds to each.
# mv .xinitrc .xinitrc.bak
v_homefile="$1"
v_filetomv="./.xinitrc"
v_logfile="/tmp/mvlog.log"

while read homedir
do
  echo "Moving to $homedir" >> "$v_logfile" 2>&1
  cd "$homedir"  >> "$v_logfile" 2>&1
  if [ -e "$v_filetomv" ];then
    echo "mving file $v_filetomv to ${v_filetomv}.bak..." >> "$v_logfile" 2>&1
    mv "${v_filetomv}" "${v_filetomv}".bak >> "$v_logfile" 2>&1
  else
    echo "$v_filetomv in $homedir not found." >> "$v_logfile" 2>&1
  fi
done < "$v_homefile"



###############################
# /etc/default/rcS
#
# Default settings for the scripts in /etc/rcS.d/
#
# For information about these variables see the rcS(5) manual page.
#
# This file belongs to the "initscripts" package.

TMPTIME=0
SULOGIN=no
DELAYLOGIN=yes
UTC=yes
VERBOSE=yes
FSCKFIX=yes



#############################
#!/bin/bash

# move to home
cd ~

# now, move down one
cd ..

echo "This will replace ALL the file in your home directory with the the ones from the "
echo "backup server.   The file servers are backed up nightly so the files will be from "
echo "the last backup.  This may not be what you want, so be VERY SURE before you "
echo "answer 'yes'!"



echo "Are you sure you want to do this? (answer with 'yes'!!)"
read ANSWER
if test  $ANSWER != "yes" 
then
	echo "You did not answer 'yes'."
	echo "good bye"
	exit 1
fi

echo "Are you REALY sure? (answer with 'yes'!!)"
read ANSWER
if test $ANSWER != "yes" 
then
	echo "You did not answer 'yes'."
	echo "good bye"
	exit 1
fi

echo "Okay here we go ...."
echo "This could take a while"

# call retrive for home directory
/usr/local/bin/retrieve -f $USER

# move back to home
cd ~

\end{verbaim}
