\section{Student and Faculty Help}\label{sec:student_faculty_help}
\subsection{Introduction}
Regardless how well structured a lab is, user support is necessary for problems and potential confusion.  This support may come in the form of a tech support email, ticketing system, website with an FAQ, or some fix utilities.  CECS labs offer a website, a tech support email, some text files with instructions, and some Linux fix scripts.
\subsection{Support Website}
The first line of support is a website.  The site should have at a minimum a frequently asked questions (FAQ) section.  FAQ pages should be very easy to find and be clearly visible.  This is the equivalent of a README file with a software distribution.  This FAQ section should answer such things as getting started, creating an account, troubleshooting accounts, resetting passwords, or answering various issues users may commonly have.  It should also have bold and/or red lettering to highlight important procedures and rules students should follow, e.g. no food or drink in the labs. 

The website is unfortunately rather pointless if users do not know where it is (or that it even exists).  In academic environments, this can somewhat be remedied by having faculty members explain to students where to go in case of a problem, or provide a handout on basic lab usage.  Of course, many students will just ignore this.  A better solution\footnote{This solution works well in non-academic environments as well because it is easier in IT department to have systems labeled with hostnames.} is to just put a label on the front of every machine with the hostname, support website, and tech email.\footnote{Tech email can be replaced with ticket helpdesk url in environments that use one.}  This will hopefully lead users to the proper place rather than coming to the technicians.  It is very important that the support website be featured prominently with clearly legible text. 

In the CECS labs, we have a support site that is the default homepage for every user.  It has everything mentioned above, however, our problem is that it does not have the latest up to the minute updates.  We do the best we can but constant changes in our environment for each release make a completely accurate website unfeasible. We strive to be accurate within a few days of the changes.

\subsection{Support Email and Alternative}

Regardless how well put together a website is, it does not account for everything.  There has to be some method of contacting tech support to report problems and ask for troubleshooting help.  The simplest way to do this is by setting up a contact email.  However, in environments with more than one technician, this contact email has to have a forwarding list.  This way, email sent will go to all techs involved.  Emails lack the ability to inform all recipients if an email has been read, or in this case, if the problem addressed in the email has been resolved.  This may cause technicians to duplicate work.  This is not much of an issue with small tech shops, especially with each tech sticking to a specialty.  But in large organizations where IT support can exceed more than a few people, support emails can be inefficient.  A solution is to set up an email with a Reply-to field so that users always answer to the single tech email that forwards to all techs.  Not all email clients properly implement this feature and some of the most common clients ignore it entirely.

Thus, a ticketing system can be a better solution.  Tickets are opened through a custom interface, e.g. a support web application.  Users login to a ticket helpdesk and open tickets describing their problem.  They may select one or more categories of problems that their issue falls under.  From there, an IT support staff can login to the helpdesk as administrators and look for tickets to work on.  A good ticketing system also has in place mechanisms for leaving notes and attachments\footnote{Attachment capability is especially necessary for screen shots.} on how the problem was resolved and what the problem was.  More advanced features might include an interface for managers to login to the helpdesk and approve requests that require management approval.\footnote{Examples of actions needing management approval include production system updates, access management changes, and operating system configuration changes.}  Lastly, the ticketing system is easier to audit compared to long email trails and is practically a necessity in environments where strict compliance is needed.   

Ticket helpdesk systems usually do not work in small environments, especially academic settings because the system is often too confusing for students to bother with.  They will usually just ignore it and continue on despite a potential problem.  Faculty will most likely never use the system and will bypass the system by emailing the first technical support contact they can.  Thus, a ticket helpdesk is probably not feasible in our environment unless required as a campus wide regulation.  

\subsection{Fix Utilities}

Sometimes a lab load cannot handle certain configurations the students and faculty may impose locally.  Various configuration files may exist in their home directories and may override system defaults.  If the configurations are not set up correctly, various issues may occur.  Sometimes these configurations may not even be intentional.  Home directories may be migrated from one distribution to another and configurations may not be compatible.  Thus, a set of fix scripts is necessary to reset configurations to a default structure.  These scripts may be located on all lab machines, but updating them would require Secure Shell (SSH) pushes to keep machines consistent even with some other repository software.  If network mounted home directories are used, it is wise to create one or more utility users (for example, ``conf") whose files are left world readable and executable.  For example, students may have a faulty ``.gnome2" configuration that could cause the gnome window manager to malfunction.  A script called ``resetX" exists in conf's home directory ``~conf/" and is world executable.  Students can run ~conf/resetX from a machine that has conf's home directory mounted or automounted.  This script will delete a student's copy of the gnome configuration and pull in a working configuration.  The script could also have logic to determine if a student has a common problem in their configuration and correct the problem without overwriting the entire configuration.  A copy of this script can be found in Appendix~\ref{ap:lab_maintenance}. 

There are a variety of other useful scripts that could be put in conf's home.  For example, you could create a script that would restore all configuration files from a tested skeleton account.  Sometimes, NFS files become locked and students do not know how to unlock them.  A script to find and unlock the files would be very helpful.  Backup and restore scripts could be useful, especially if a nightly backup system is run.  Care should be taken as to where these scripts should reside, however.  If the script often changes it should be put in ~conf/ and shared over the network.  If the script is somewhat more permanent, it is often easier to just house the script locally.  The recommended location for these scripts is in ``/usr/local/bin/" where they are made world executable.  Thus, with each master image revision they will be updated if needed.  
